equations

### 2023-05-22 equations.tex
updates

---
title: "equations"
author: "Jesper Fischer Ehmsen"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The Rescorla wagner model is a mathematical model of how classical conditions are thought to occur. The basic idea is that an agent updates his belief based on a weighted prediction error which is the difference between the last cue-stimulus association (association) and the last belief. The weighing of the prediction error is generally called the learning rate as this parameter determines the degree to which an agent utilises the new information to update his belief. This learning rate $\alpha$ is for the Rescorla Wagner agent assumed to be constant.

$$
\begin{equation}
\tag{1}
E_t = E_{t-1}+\alpha*\underbrace{(A_{t-1}-E_{t-1})}_{\text{Prediction Error (PE)}}
\end{equation}
$$
In the current experiment we modify this update equation to encapsulate the interaction between precept and expectations. We do this by modifying how the prediction error is calculated as now the prediction error is going to be calculated from the precept  $\psi_t$ that the expectation and the stimulus elicits. We assume for the RW agent that the weighting of stimulus and expectation is governed by a single parameter $\gamma$, like the learning rate $\alpha$ it determines the weight that is put on the stimulus vs expectation in determining what the participant felt on the current trial $\psi_t$

$$
\begin{equation}
\tag{2}
\psi_t = \gamma*S_t+(1-\gamma)*E_t
\end{equation}
$$

$$
\begin{equation}
\tag{3}
PE_t = \psi_t-E_{t-1}
\end{equation}
$$

#### weighted bayes

@@

$$
\begin{equation}
\tag{4}
\psi_t = S(w_1*S^{-1}(S_t)+w_2*S^{-1}(E_1))
\end{equation}
$$
where $S(x) = \frac{1}{1+e^{-x}}$. The rest of the update equations remain similar to the Resorla Wagner, essentially also utilizing the RW model to update the expectation on each trial with the prediction error being (3)

##### Kalman filter

@@

$$
\begin{equation}
\tag{5}
S_t \sim \mathcal{N}{(\psi{_t}, \sigma_s^2)}
\end{equation}
$$

@@

$$
\begin{equation}
\tag{6}
\psi_t \sim \mathcal{N}{(E_t, \sigma_{\psi}^2)}
\end{equation}
$$
Where this mean is updated based on the stimulus given the cue on previous trials

$$
\begin{equation}
\tag{7}
E_t \sim \mathcal{N}{(E_{t-1}, \sigma_\eta^2)}
\end{equation}
$$
Lastly this prior expectation, i.e. the expectation of the last trial is given by all the previous inputs which the agent keeps track off.

$$
\begin{equation}
\tag{8}
E_t | S_{{t-1}} \sim \mathcal{N}{(\mu_{E_{t-1}}, \sigma_{E_t}^2)}
\end{equation}
$$
Now combining (8) and (6) we get a prior for the precept at a given trial, which is before getting the stimulus but after getting the cue

$$
\begin{equation}
\tag{9}
\psi_t | S_{{t-1}} \sim \mathcal{N}{(\mu_{E_{t-1}}, \sigma_{E_t}^2+ \sigma_{\psi}^2))}
\end{equation}
$$

@@

$$
\begin{equation}
\tag{10}
\psi_t | S_{{t}} \sim \mathcal{N}{(\frac{\sigma_{s}^2*\mu_{E_{t-1}}+(\sigma_{\psi}^2+\sigma_{E_t}^2)*S_t}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}, \frac{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}{\sigma_{s}^2*(\sigma_{\psi}^2+\sigma_{E_t}^2)}})
\end{equation}
$$
The mean rating of the agent on trial t is therefore given by the mean, which can also be writting as:

$$
\begin{equation}
\tag{11}
\mu_{\psi_t} = \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}*\mu_{E_{t-1}}+\frac{\sigma_{\psi}^2+\sigma_{E_t}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}*S_t
\end{equation}
$$

@@

$$
\begin{equation}
\tag{12}
E_{t}|S_t \sim \mathcal{N}{(\frac{(\sigma_{s}^2+\sigma_{\psi}^2)*\mu_{E_{t-1}}+\sigma_{E_t}^2*S_t}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}, \frac{(\sigma_{s}^2+\sigma_{\psi}^2)*\sigma_{E_t}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}+\sigma_\eta^2})
\end{equation}
$$

It can be shown that these update equations for the expectation and the mean percept are a generalization of the above mentioned RW model if we define two trial by trial parameters $\gamma_{t}$ and $\alpha_{t}$

$$
\begin{equation}
\tag{13}
\gamma_{t} = \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}
\end{equation}
$$
$$
\begin{equation}
\tag{14}
\alpha_{t} = \frac{\sigma_{E_t}^2}{\sigma_{\psi}^2+\sigma_{E_t}^2}
\end{equation}
$$

@@

$$
\begin{equation}
\tag{15}
\psi_{t} = (1-\gamma_{t})*S_{t}+\gamma_{t}*E_{t}
\end{equation}
$$

$$
\begin{equation}
\tag{16}
E_{t} = \alpha_{t}*\psi_{t}+(1-\alpha_{t})*E_{t}
\end{equation}
$$
For the derivation of equation 15 and 16 readers are refereed to the supplementary material.

```{r}

```

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={equations},
  pdfauthor={Jesper Fischer Ehmsen},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{equations}
\author{Jesper Fischer Ehmsen}
\date{2023-05-22}

\begin{document}
\maketitle

\hypertarget{r-markdown}{%
\subsection{R Markdown}\label{r-markdown}}

The Rescorla wagner model is a mathematical model of how classical
conditions are thought to occur. The basic idea is that an agent updates
his belief based on a weighted prediction error which is the difference
between the last cue-stimulus association (association) and the last
belief. The weighing of the prediction error is generally called the
learning rate as this parameter determines the degree to which an agent
utilises the new information to update his belief. This learning rate
\(\alpha\) is for the Rescorla Wagner agent assumed to be constant.

\[
\begin{equation}
\tag{1}
E_t = E_{t-1}+\alpha*\underbrace{(A_{t-1}-E_{t-1})}_{\text{Prediction Error (PE)}}
\end{equation}
\] In the current experiment we modify this update equation to
encapsulate the interaction between precept and expectations. We do this
by modifying how the prediction error is calculated as now the
prediction error is going to be calculated from the precept \(\psi_t\)
that the expectation and the stimulus elicits. We assume for the RW
agent that the weighting of stimulus and expectation is governed by a
single parameter \(\gamma\), like the learning rate \(\alpha\) it
determines the weight that is put on the stimulus vs expectation in
determining what the participant felt on the current trial \(\psi_t\)

\[
\begin{equation}
\tag{2}
\psi_t = \gamma*S_t+(1-\gamma)*E_t
\end{equation}
\]

\[
\begin{equation}
\tag{3}
PE_t = \psi_t-E_{t-1}
\end{equation}
\]

\hypertarget{weighted-bayes}{%
\paragraph{weighted bayes}\label{weighted-bayes}}

@@

\[
\begin{equation}
\tag{4}
\psi_t = S(w_1*S^{-1}(S_t)+w_2*S^{-1}(E_1))
\end{equation}
\] where \(S(x) = \frac{1}{1+e^{-x}}\). The rest of the update equations
remain similar to the Resorla Wagner, essentially also utilizing the RW
model to update the expectation on each trial with the prediction error
being (3)

\hypertarget{kalman-filter}{%
\subparagraph{Kalman filter}\label{kalman-filter}}

@@

\[
\begin{equation}
\tag{5}
S_t \sim \mathcal{N}{(\psi{_t}, \sigma_s^2)}
\end{equation}
\]

@@

\[
\begin{equation}
\tag{6}
\psi_t \sim \mathcal{N}{(E_t, \sigma_{\psi}^2)}
\end{equation}
\] Where this mean is updated based on the stimulus given the cue on
previous trials

\[
\begin{equation}
\tag{7}
E_t \sim \mathcal{N}{(E_{t-1}, \sigma_\eta^2)}
\end{equation}
\] Lastly this prior expectation, i.e.~the expectation of the last trial
is given by all the previous inputs which the agent keeps track off.

\[
\begin{equation}
\tag{8}
E_t | S_{{t-1}} \sim \mathcal{N}{(\mu_{E_{t-1}}, \sigma_{E_t}^2)}
\end{equation}
\] Now combining (8) and (6) we get a prior for the precept at a given
trial, which is before getting the stimulus but after getting the cue

\[
\begin{equation}
\tag{9}
\psi_t | S_{{t-1}} \sim \mathcal{N}{(\mu_{E_{t-1}}, \sigma_{E_t}^2+ \sigma_{\psi}^2))}
\end{equation}
\]

@@

\[
\begin{equation}
\tag{10}
\psi_t | S_{{t}} \sim \mathcal{N}{(\frac{\sigma_{s}^2*\mu_{E_{t-1}}+(\sigma_{\psi}^2+\sigma_{E_t}^2)*S_t}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}, \frac{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}{\sigma_{s}^2*(\sigma_{\psi}^2+\sigma_{E_t}^2)}})
\end{equation}
\] The mean rating of the agent on trial t is therefore given by the
mean, which can also be writting as:

\[
\begin{equation}
\tag{11}
\mu_{\psi_t} = \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}*\mu_{E_{t-1}}+\frac{\sigma_{\psi}^2+\sigma_{E_t}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}*S_t
\end{equation}
\]

@@

\[
\begin{equation}
\tag{12}
E_{t}|S_t \sim \mathcal{N}{(\frac{(\sigma_{s}^2+\sigma_{\psi}^2)*\mu_{E_{t-1}}+\sigma_{E_t}^2*S_t}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}, \frac{(\sigma_{s}^2+\sigma_{\psi}^2)*\sigma_{E_t}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}+\sigma_\eta^2})
\end{equation}
\]

It can be shown that these update equations for the expectation and the
mean percept are a generalization of the above mentioned RW model if we
define two trial by trial parameters \(\gamma_{t}\) and \(\alpha_{t}\)

\[
\begin{equation}
\tag{13}
\gamma_{t} = \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}
\end{equation}
\] \[
\begin{equation}
\tag{14}
\alpha_{t} = \frac{\sigma_{E_t}^2}{\sigma_{\psi}^2+\sigma_{E_t}^2}
\end{equation}
\]

@@

\[
\begin{equation}
\tag{15}
\psi_{t} = (1-\gamma_{t})*S_{t}+\gamma_{t}*E_{t}
\end{equation}
\]

\[
\begin{equation}
\tag{16}
E_{t} = \alpha_{t}*\psi_{t}+(1-\alpha_{t})*E_{t}
\end{equation}
\] For the derivation of equation 15 and 16 readers are refereed to the
supplementary material.

\end{document}
